---
title: "Permits and ED"
author: "ignacio"
date: "March 30, 2017"
output: github_document
html_document:
  df_print: paged
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )

setwd("C:/Users/icps/Dropbox/3. Maxwell/3. Spring Term/5. Data Driven II/Labs/all-labs-ddmii-icps86/Permits_and_EC/")
dir()

dat <- read.csv("C:/Users/icps/Dropbox/3. Maxwell/3. Spring Term/5. Data Driven II/Labs/Instructions/Permits_ED/PrivateInvestment_Clean.csv", stringsAsFactors = F)

library(pander)
library(dplyr)


```
###Looking at the Permit Data set
```{r}
types <- data.frame(type = names(table(dat$Type)), freq = as.numeric(table(dat$Type)))
types <- types[order(types$freq, decreasing = T),]
rownames(types) <- NULL
pander(types)
```


**Residential**

* Res. Remodel/Chg Occ :  Residential remodeling / Change in occupanc?
* Res. New 1-2 Family

**businesses**

* Loading Zone (Business)
* Block Party (Business)
* Parking Meter Rental

**uncertain** - difficult to tell if its residential or business

* Com. New Building        
* Demolition              
* Site Work
* Footing / Foundation

**Not sure if should be used**

* HVAC/Mechanical:        Heating, ventilation, AC or mechanical repair. (uncertain)
* Com. Reno/Rem/Chg Occ:  ???????
* Electric (Meter Set)

**to be Ignored**

* Sprinkler
* Misc.(deck, fence,ramp) 
* Sidewalk Replace
* Fire Alarm
* Electric
* Liability Waiver
* Road Cut                 
* Security Alarm
* Sign
* Sidewalk Cafe            
* Elevator                 
* Public Assembly
* Block Party (Residential)
* Antenna / Dish
* Pool / Hot Tub           
* Tank
* Curb Cut
* Encroach (Major)
* Encroachment (Converted)
* Encroach (Deminimus)

```{r}
types$type <- as.character(types$type)
class(types$type)

#x <- types$type == "Res. Remodel/Chg Occ" | "Res. New 1-2 Family" | "Loading Zone (Business)" | "Block Party (Business)" | "Parking Meter Rental" | "Com. New Building" | "Demolition" | "Site Work" | "Footing / Foundation" | "HVAC/Mechanical" | "Com. Reno/Rem/Chg Occ" | "Electric (Meter Set)"

```


