---
title: "Permits"
author: "ignacio"
date: "March 30, 2017"
output: github_document
html_document:
  df_print: paged
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F , fig.width = 10, fig.height = 10)

setwd("C:/Users/icps/Dropbox/3. Maxwell/3. Spring Term/5. Data Driven II/Labs/all-labs-ddmii-icps86/Permits_and_EC/")

library(dplyr)
library(pander)
library(ggmap)
library(sp) #coordinates function ir here
#install.packages("geojsonio")
library(geojsonio )
library(maptools )
library(rgdal)
library(spatialEco)

```

```{r, echo=FALSE}

#LOADING DATASET, formating it for sp and creating a types subset

dat <- read.csv("https://raw.githubusercontent.com/R-Class/all-labs-ddmii-icps86/master/Permits_and_EC/Permits_processed.csv", stringsAsFactors = F)

#Creating two subsets: dat and df.Ign

#1. df.Ign subset holds  the permit types we are not using.
#creating a vector that contains all the cases of ignored permit types.
x <- dat$Type
Ign <- x == "Antenna / Dish" | x =="Electric (Meter Set)" |  x=="Block Party (Business)" | x=="Block Party (Residential)" | x=="Curb Cut" | x=="Encroach (Deminimus)" | x=="Encroach (Major)"  | x=="Encroachment (Converted)" | x=="Footing / Foundation"  | x=="Liability Waiver" | x=="Loading Zone (Business)" | x=="Parking Meter Rental" | x=="Public Assembly" | x=="Road Cut" | x=="Sidewalk Cafe" | x=="Sidewalk Replace" | x=="Sign" | x=="Site Work"

#subsetting it (18 ignored permit types)
df.Ign <- dat[Ign,]

#2. dat subset holds all permit types we are going to use. These are geocoded and have tract id (14 permit types)
dat <- dat[!Ign,]

#CLIPPING all dat permits outside syracuse city
#use the tract variable, because of the over spatial function (described in the data wrangling rmd) the points outside the tracts have NA

x<- is.na(dat$Tract) #191 NAs, or points outside the tracts
dat <- dat[!x,] # eliminating the NAs

####################

#Creating the Dataframe with aggregates for each variable

#TOTAL 
x <- group_by(dat, Tract, Year)
total <- summarize(x, 
                   Tot.Frq = n(),
                   Tot.Val = sum(Valuation)
                   )

#df.R
cat <- as.character(dat$Type)  #setting up indexes
df.R <- cat== "Res. Remodel/Chg Occ" | cat == "Res. New 1-2 Family" 
df.R <- dat[df.R,] 

x <- group_by(df.R, Tract, Year)
Res <- summarize(x, 
                   Res.Frq = n(),
                   Res.Val = sum(Valuation)
                   )

#df.C
df.C <- cat== "Com. New Building" | cat=="Com. Reno/Rem/Chg Occ"
df.C <- dat[df.C,] 

x <- group_by(df.C, Tract, Year)
Com <- summarize(x, 
                   Com.Frq = n(),
                   Com.Val = sum(Valuation)
                   )

#df.D
df.D <- cat=="Demolition"
df.D <- dat[df.D,] 

x <- group_by(df.D, Tract, Year)
Dem <- summarize(x, 
                   Dem.Frq = n(),
                   Dem.Val = sum(Valuation)
                   )

#df.I
df.I <- cat=="Electric" | cat=="Elevator" | cat=="Fire Alarm" | cat=="HVAC/Mechanical" | cat=="Misc.(deck, fence,ramp)" | cat=="Pool / Hot Tub" | cat=="Security Alarm" | cat=="Sprinkler" | cat=="Tank"
df.I <- dat[df.I,] 

x <- group_by(df.I, Tract, Year)
Ins <- summarize(x, 
                   Ins.Frq = n(),
                   Ins.Val = sum(Valuation)
                   )

#binding the different aggregates by variable
var <- merge(total, Res, all = T)
var <- merge(var,Com, all = T)
var <- merge(var,Ins, all = T)

#write.csv(var, file = "Permits_var.csv")


####################

#FORMATTING dat as a spatial object
dat <- SpatialPointsDataFrame(dat[ ,c( "lon", "lat") ], dat, proj4string=CRS("+proj=longlat +datum=WGS84")) #this is so that it does not lose the lat lon in the dataframe when transformed into a sp


####################

#CREATING types SUBSET for dat and df.Ign: just to show the freq of the different permits types

#First for the dat dataset.
#using table to create a dataframe with the frequencies for each category
types <- data.frame(type = names(table(dat$Type)), freq = as.numeric(table(dat$Type)))
rownames(types) <- NULL
types <- arrange(types, type) #ordering types alphabetically

#getting the mean/max/min value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=mean)
x <- arrange(x, Group.1)
types$meanval <- x$x

#getting the min value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=min)
x <- arrange(x, Group.1)
types$minval <- x$x

#getting the max value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=max)
x <- arrange(x, Group.1)
types$maxval <- x$x

#Second, for the df.Ign dataset
#using table to create a dataframe with the frequencies for each category
types.Ign <- data.frame(type = names(table(df.Ign$Type)), freq = as.numeric(table(df.Ign$Type)))
rownames(types.Ign) <- NULL
types.Ign <- arrange(types.Ign, type) #ordering types alphabetically

#getting the mean/max/min value per type
x <- aggregate(df.Ign$Valuation, by=list(df.Ign$Type), FUN=mean)
x <- arrange(x, Group.1)
types.Ign$meanval <- x$x

#getting the min value per type
x <- aggregate(df.Ign$Valuation, by=list(df.Ign$Type), FUN=min)
x <- arrange(x, Group.1)
types.Ign$minval <- x$x

#getting the max value per type
x <- aggregate(df.Ign$Valuation, by=list(df.Ign$Type), FUN=max)
x <- arrange(x, Group.1)
types.Ign$maxval <- x$x


#########################

#LOADING THE SHAPEFILE

#Loading a geojson file of only census tracts
shapes <- geojson_read("https://raw.githubusercontent.com/R-Class/all-labs-ddmii-icps86/master/Permits_and_EC/tracts.geojson", method="local", what="sp" )


```



##Information about the Dataset
The Permits Dataset has information for 21,556 permits from 2012 - 2013. 
For each permit the Dataset inclueds variables like:

* Type of Permit
* Applicant name (can be either a person or a company)
* Location
* SBL
* Date Issued
* Valuation (cost of the activity the applicant is requesting permit for)
* Fee Amount (the monies for the City)

We used the types of permits to construct 4 categories that we will aggregate as variables at the Census Tract level.

There are 32 types of permits:

We have placed this types of Permits into 4 categories to analyze them separately. 
They are:

* Residential Properties
* Commercial Properties
* Demolitions
* Installations and Repairs (to public space, res and com)


In what follows we explain the categories that we have constructed and show some descriptives.

###PROPOSED VARIABLES

####**1.New/Reno of Residential Property**

```{r, echo=FALSE}
#types. creating an index of the types to run the following code chunks
x <- as.character(types$type) #setting up indexes
Res <- x== "Res. Remodel/Chg Occ" | x == "Res. New 1-2 Family" 
Res <- types[Res,]
pander(Res)

#df.R
cat <- as.character(dat@data$Type)  #setting up indexes
df.R <- cat== "Res. Remodel/Chg Occ" | cat == "Res. New 1-2 Family" 
df.R <- dat[df.R,] 

pander(table(df.R$Year))
summary(df.R$Valuation) %>% pander
barplot(sort(df.R$Valuation, decreasing = T))

#plotting Residential 
plot(shapes, col = F, border = "grey10") 
  title(main = "Residential Permits")
  points(df.R, pch=20, cex = .4, col=adjustcolor("dodgerblue4", alpha.f = .8))

```

###**2.New/Reno of Commercial Property**
```{r, echo=FALSE}
#types
Com <- x== "Com. New Building" | x=="Com. Reno/Rem/Chg Occ"
Com <- types[Com,]
pander(Com)

#df
df.C <- cat== "Com. New Building" | cat=="Com. Reno/Rem/Chg Occ"
df.C <- dat[df.C,] 

pander(table(df.C$Year))
summary(df.C$Valuation) %>% pander
barplot(sort(df.C$Valuation, decreasing = T))

plot(shapes, col = F, border = "grey10") 
  title(main = "Commercial Permits")
  points(df.C, pch=20, cex= .4, col=adjustcolor("red2", alpha.f = .8))
  
```


###**3.Demolitions**
```{r, echo=FALSE}
#types
Dem <- x=="Demolition"
Dem <- types[Dem,]
pander(Dem)

#df
df.D <- cat=="Demolition"
df.D <- dat[df.D,] 

pander(table(df.D$Year))
summary(df.D$Valuation) %>% pander
barplot(sort(df.D$Valuation, decreasing = T))

plot(shapes, col = F, border = "grey10")  
  title(main = "Demolition Permits")
  points(df.D, pch=20, cex= .4, col=adjustcolor("forestgreen", alpha.f = .8))

```

###**4.Installations and Repairs on all properties**

*At the end of this page each type of permit that makes up this category is plotted. 

```{r, echo=FALSE}
#types
Ins <- x=="Electric" | x=="Elevator" | x=="Fire Alarm" | x=="HVAC/Mechanical" | x=="Misc.(deck, fence,ramp)" | x=="Pool / Hot Tub" | x=="Security Alarm" | x=="Sprinkler" | x=="Tank"
Ins <- types[Ins,]
pander(Ins)

#df
df.I <- cat=="Electric" | cat=="Elevator" | cat=="Fire Alarm" | cat=="HVAC/Mechanical" | cat=="Misc.(deck, fence,ramp)" | cat=="Pool / Hot Tub" | cat=="Security Alarm" | cat=="Sprinkler" | cat=="Tank"
df.I <- dat[df.I,] 

pander(table(df.I$Year))
summary(df.I$Valuation) %>% pander
barplot(sort(df.I$Valuation, decreasing = T))

plot(shapes, col = F, border = "grey10") 
  title(main = "Instalations and Repairs")
  points(df.I, pch=20, cex= .4, col=adjustcolor("orange", alpha.f = .6))

```


###All Categories at a glance

```{r, echo=FALSE}
par( mar = c( 1, 1, 3, 1 ) )
plot(shapes, col = F, border = "grey10") 
  title(main = "All categories of Permits")
  points(df.I, pch=20, cex= .4, col=adjustcolor("orange", alpha.f = .8))
  points(df.D, pch=20, cex= .4, col=adjustcolor("forestgreen", alpha.f = .8))
  points(df.C, pch=20, cex= .4, col=adjustcolor("red2", alpha.f = .8))
  points(df.R, pch=20, cex = .4, col=adjustcolor("dodgerblue4", alpha.f = .8))
  legend("bottomright", legend = c("Residential", "Commercial", "Insta/Rep", "Demolitions"), bty = "n", pch = 20, cex = .8, pt.cex = 2, col = c("orange","forestgreen","red2","dodgerblue4"))

```



```{r,echo=FALSE}
#EXPLORING THE TYPES IN THE INSTALLATION CATEGORY

#Electric
  x <- df.I$Type == "Electric"
plot(shapes) 
  title(main= "Electric Permits")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

#Elevator
x <- df.I$Type == "Elevator"
plot(shapes) 
  title(main= "Elevator")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

#Fire Alarm
x <- df.I$Type == "Fire Alarm"
plot(shapes) 
  title(main= "Fire Alarm installations")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

#HVAC/mech
  x <- df.I$Type == "HVAC/Mechanical"
plot(shapes) 
  title(main= "HVAX/MEch (AirConditionin and heating")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

#Misc.(deck, fence,ramp) 
x <- df.I$Type == "Misc.(deck, fence,ramp)"
plot(shapes) 
  title(main= "Misc. (deck, fence, ramp)")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

# Pool / Hot Tub
x <- df.I$Type == "Pool / Hot Tub"
plot(shapes) 
  title(main= "Hot tub")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

# Security alarm
x <- df.I$Type == "Security Alarm"
plot(shapes) 
title(main= "Security Alarm")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

# Sprinkler 
  x <- df.I$Type == "Sprinkler"
plot(shapes) 
  title(main= "Sprinkler system")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

# Tank
  x <- df.I$Type == "Tank"
plot(shapes) 
  title(main= "Tank")
  points(df.I[x,], pch=20, cex= .2, col=adjustcolor("red", alpha.f = .6))

    
```





###**Categories not taken into account**

All permits we found it was not useful to include:

```{r, echo=FALSE}
pander(types.Ign)

#the df.Ign already exists
pander(table(df.Ign$Year))
paste("total cases:", nrow(df.Ign))
summary(df.Ign$Valuation) %>% pander
barplot(sort(df.Ign$Valuation, decreasing = T))

```

####**Why is New Foundations not included?**

"Footing / foundation" type of permits are not being included because it is only 16 and it is not clear whether they overlap with the permits for new residential and/or new commercial. In other words, a new building will have a foundaiton permit and a new construction permit (either residential or commercial).

```{r, echo=FALSE}
Found <- types.Ign$type == "Footing / Foundation"
pander(types.Ign[Found, ])

#df
Found <- df.Ign$Type == "Footing / Foundation"
df.F <- df.Ign[Found,] 

pander(table(df.F$Year))
summary(df.F$Valuation) %>% pander
barplot(sort(df.F$Valuation, decreasing = T))

df.F <- SpatialPointsDataFrame(df.F[ ,c( "lon", "lat") ], df.F, proj4string=CRS("+proj=longlat +datum=WGS84")) #this is so that it does not lose the lat lon in the dataframe when transformed into a sp


plot(shapes, col = F, border = "grey10")  
  title(main = "Foundation starts Permits")
  points(df.F, pch=20, cex = 1.2, col=adjustcolor("deeppink4", alpha.f = 1))


```
