---
title: "Permits and ED"
author: "ignacio"
date: "March 30, 2017"
output: github_document
html_document:
  df_print: paged
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F )

setwd("C:/Users/icps/Dropbox/3. Maxwell/3. Spring Term/5. Data Driven II/Labs/all-labs-ddmii-icps86/Permits_and_EC/")

library(pander)
library(dplyr)
```


```{r, echo=FALSE, eval=FALSE}
#Formatting the dataframe
dat <- read.csv("C:/Users/icps/Dropbox/3. Maxwell/3. Spring Term/5. Data Driven II/Labs/Instructions/Permits_ED/PrivateInvestment_Clean.csv", stringsAsFactors = F)

#making valuation and fee variables to be numeric
x <- dat$Valuation
x <- gsub("$", "", x, fixed = T)
x <- gsub(",", "", x, fixed = T)
x <- as.numeric(x)
dat$Valuation <- x

x <- dat$Fee.Amount
x <- gsub("$", "", x, fixed = T)
x <- gsub(",", "", x, fixed = T)
x <- as.numeric(x)
dat$Fee.Amount <- x

save(dat, file= "Permits.rda")

```

```{r, echo=FALSE, eval=FALSE}
#Generating a dataframe about the permits types

#using table to create a dataframe with the frequencies for each category
types <- data.frame(type = names(table(dat$Type)), freq = as.numeric(table(dat$Type)))
types <- types[order(types$freq, decreasing = T),]
rownames(types) <- NULL
types <- arrange(types, type) #ordering types alphabetically

#getting the mean/max/min value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=mean)
x <- arrange(x, Group.1)
types$meanval <- x$x

#getting the min value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=min)
x <- arrange(x, Group.1)
types$minval <- x$x

#getting the max value per type
x <- aggregate(dat$Valuation, by=list(dat$Type), FUN=max)
x <- arrange(x, Group.1)
types$maxval <- x$x

save(types, file= "types.rda")

```

#SOME INFORMATION

###CENSUS NEW RESIDENTIAL CONSTRUCTION REPORT

The US Census Bureau and HUD release a MONTHLY NEW RESIDENTIAL CONSTRUCTION REPORT where they show 3 indicators:

The Privately-owned:
- Housing permits 
- Housing starts
- Housing completion

(from: https://www.census.gov/construction/nrc/index.html)


###Links

Business improvement districts
https://business.syrgov.net/open/business-improvement-districts-bids/

Vaccant property registry
http://www.syracuse.ny.us/VacantRegistry.aspx 


###PROPOSED INDICATORS

```{r, echo=FALSE}
load("Permits.rda")
load("types.rda")

#CREATING THE SUBSETS OF TYPES TABLE TO SHOW SUMMARY DESCRIPTIVES OF PERMITS IN EACH VARIABLE
x<- as.character(types$type)

#Foundation
Found <- x == "Footing / Foundation"
Found <- types[Found, ]

#Res
Res <- x== "Res. Remodel/Chg Occ" | x == "Res. New 1-2 Family" 
Res <- types[Res,]

#Com
Com <- x== "Com. New Building" | x=="Com. Reno/Rem/Chg Occ"
Com <- types[Com,]

#Dem
Dem <- x=="Demolition"
Dem <- types[Dem,]

#Ins
Ins <- x=="Antenna / Dish" | x=="Electric" | x=="Electric (Meter Set)" | x=="Elevator" | x=="Fire Alarm" | x=="HVAC/Mechanical" | x=="Misc.(deck, fence,ramp)" | x=="Pool / Hot Tub" | x=="Security Alarm" | x=="Sprinkler" | x=="Tank"
Ins <- types[Ins,]


#Ign
Ign <- x=="Block Party (Business)" | x=="Block Party (Residential)" | x=="Curb Cut" | x=="Encroach (Deminimus)" | x=="Encroach (Major)"  | x=="Encroachment (Converted)" | x=="Liability Waiver" | x=="Loading Zone (Business)" | x=="Parking Meter Rental" | x=="Public Assembly" | x=="Road Cut" | x=="Sidewalk Cafe" | x=="Sidewalk Replace" | x=="Sign" | x=="Site Work"
Ign <- types[Ign,]  
  
```

```{r, echo=FALSE, eval=FALSE}
#CREATING DF SUBSETS from de dat Dataframe FOR EACH VARIABLE

#Found
x <- as.character(dat$Type)

df.F <- x == "Footing / Foundation"
df.F <- dat[df.F, ]

#Res
df.R <- x== "Res. Remodel/Chg Occ" | x == "Res. New 1-2 Family" 
df.R <- dat[df.R,]

#Com
df.C <- x== "Com. New Building" | x=="Com. Reno/Rem/Chg Occ"
df.C <- dat[df.C,]

#Dem
df.D <- x=="Demolition"
df.D <- dat[df.D,]

#Ins
df.I <- x=="Antenna / Dish" | x=="Electric" | x=="Electric (Meter Set)" | x=="Elevator" | x=="Fire Alarm" | x=="HVAC/Mechanical" | x=="Misc.(deck, fence,ramp)" | x=="Pool / Hot Tub" | x=="Security Alarm" | x=="Sprinkler" | x=="Tank"
df.I <- dat[df.I,]


#Ign
df.Ign <- x=="Block Party (Business)" | x=="Block Party (Residential)" | x=="Curb Cut" | x=="Encroach (Deminimus)" | x=="Encroach (Major)"  | x=="Encroachment (Converted)" | x=="Liability Waiver" | x=="Loading Zone (Business)" | x=="Parking Meter Rental" | x=="Public Assembly" | x=="Road Cut" | x=="Sidewalk Cafe" | x=="Sidewalk Replace" | x=="Sign" | x=="Site Work"
df.Ign <- dat[df.Ign,]  

```


```{r, echo=FALSE, eval=FALSE}
#geocoding

library(ggmap)
#devtools::install_github("dkahle/ggmap") #Installing latest ggmap from the creators webpage. Need to download this in order for the code to work.

register_google(key = "YOUR KEY HERE", account_type = "premium", day_limit = 100000) #Need to put premium to fool the function. I did not have a premium, just a google API key that was authorized to bill me
ggmap_credentials()
geocodeQueryCheck()

#in what follows I geocoded 17,933 permits in approx 3h

#Found
df.F$Location2 <- paste(df.F$Location, ", Syracuse, New York")
df.F <- mutate_geocode(df.F, Location2, source = "google")
save(df.F, file = "df.F.rda")

#Res
df.R$Location2 <- paste(df.R$Location, ", Syracuse, New York")
df.R <- mutate_geocode(df.R, Location2, source = "google")
save(df.R, file = "df.R.rda")

#Com
df.C$Location2 <- paste(df.C$Location, ", Syracuse, New York")
df.C <- mutate_geocode(df.C, Location2, source = "google")
save(df.C, file = "df.C.rda")

#Dem
df.D$Location2 <- paste(df.D$Location, ", Syracuse, New York")
df.D <- mutate_geocode(df.D, Location2, source = "google")
save(df.D, file = "df.D.rda")

#Ins
df.I$Location2 <- paste(df.I$Location, ", Syracuse, New York")
df.I <- mutate_geocode(df.I, Location2, source = "google")
save(df.I, file = "df.I.rda")

save(df.Ign, file = "df.Ign.rda")
```

```{r, echo=FALSE}
#loading geocoded df
load("df.C.rda")
load("df.D.rda")
load("df.F.rda")
load("df.I.rda")
load("df.R.rda")
load("df.Ign.rda")
```

**1.New Foundations starts (Found)**
```{r, echo=FALSE}
pander(Found)
table(is.na(df.F$lon)) %>% pander
summary(df.F$Valuation) %>% pander
barplot(sort(df.F$Valuation, decreasing = T))

```

**2.New/Reno Residential (Res)**
```{r, echo=FALSE}
pander(Res)
table(is.na(df.R$lon)) %>% pander
summary(df.R$Valuation) %>% pander
barplot(sort(df.R$Valuation, decreasing = T))

```

**3.New/Reno Commercial (Com)**
```{r, echo=FALSE}
pander(Com)
table(is.na(df.C$lon)) %>% pander
summary(df.C$Valuation) %>% pander
barplot(sort(df.C$Valuation, decreasing = T))

```

**4.Demolitions (Dem)**
```{r, echo=FALSE}
pander(Dem)
table(is.na(df.D$lon)) %>% pander
summary(df.D$Valuation) %>% pander
barplot(sort(df.D$Valuation, decreasing = T))

```

**5.Installations and Repairs (Ins)**
```{r, echo=FALSE}
pander(Ins)
table(is.na(df.I$lon)) %>% pander
summary(df.I$Valuation) %>% pander
barplot(sort(df.I$Valuation, decreasing = T))
```

**Ignored (Ign)**
```{r, echo=FALSE}
pander(Ign)
paste("total cases:", nrow(df.Ign))
summary(df.Ign$Valuation) %>% pander
barplot(sort(df.Ign$Valuation, decreasing = T))

```


